---
title: "Chapter 2: Statistical Learning"
output: html_notebook
---

## 8. College Dataset

### (a) load dataset
Datasets available on http://www-bcf.usc.edu/~gareth/ISL/data.html

```{r}
# read the data in csv format from the web
College = read.csv('http://www-bcf.usc.edu/~gareth/ISL/College.csv')

# display the table
head(College)
```

### (b) assign the 1st column to rownames
```{r}
# assign the rownames to the 1st column
rownames(College) = College[,1]

# remove the 1st column that is now duplicate
College = College[,-1]

# display the table
head(College)
```

### (c) inspect the data
#### i. see summary
```{r}
summary(College)
```

#### ii. plot scatterplot
```{r}
# plot just the first 10 entries
pairs(College[,1:10])

#check correlations
library(corrr)
library(magrittr)
College[2:18] %>% correlate() %>% rearrange() %>% shave() %>% rplot()
```

#### iii. boxplot of `Outstate` vs `Private`
```{r}
plot(College$Private, College$Outstate, xlab='Private', ylab='Outstate Applications')
```

#### iv. create `Elite` qualitative variable
```{r}
Elite = rep("No", nrow(College))
Elite[College$Top10perc>50] = "Yes"
Elite=as.factor(Elite)
College=data.frame(College, Elite)

summary(College$Elite)
plot(College$Elite, College$Outstate, xlab='Elite', ylab='Outstate Applications')

```
#### v. plot histograms
```{r}
par(mfrow=c(2,2))
hist(College$Apps, breaks=20, main="", xlab="Applications")
hist(College$Accept, breaks=20, main="", xlab="Acceptance")
hist(College$Enroll, breaks=20, main="", xlab="Enrolled")
hist(College$Top10perc, breaks=20, main="", xlab="Top 10%")
```

## 9. Auto Dataset
```{r}
# load dataset
Auto = read.csv('http://www-bcf.usc.edu/~gareth/ISL/Auto.csv', header=T, na.strings="?")

# remove missing values
Auto = na.omit(Auto)

# fix horsepower being seen as categorical
Auto$horsepower = as.integer(Auto$horsepower)

# make origin categorical
Auto$origin = as.factor(Auto$origin)

head(Auto)
```

### (a) qualitative and quantitative predictors
```{r}
summary(Auto)
```

* Quantitative: mpg, cylinders, displacement, weight, accelerations, year, origin
* Qualitative: origin, name

### (b) range for each quantitative predictor
```{r}
# sapply range only to quantitative columns
sapply(Auto[, 1:7], range)
```

### (c) mean and standard deviation for each quantitative predictor
```{r}
# mean
sapply(Auto[, 1:7], mean)

# standard deviation
sapply(Auto[, 1:7], sd)
```

### (d) remove 10th through 85th observations and calculate range mean and standard deviation
```{r}
# remove rows
Auto_subset = Auto[-(10:85),]

# range
sapply(Auto_subset[, 1:7], range)

# mean
sapply(Auto_subset[, 1:7], mean)

# standard deviation
sapply(Auto_subset[, 1:7], sd)

```

### (e) analyse data visually
```{r}
pairs(Auto)

plot(Auto$horsepower, Auto$mpg, xlab="horse power", ylab="mpg")

plot(as.factor(Auto$cylinders), Auto$mpg, xlab="cylinders", ylab="mpg")
```

### (f) predict `mpg` from other variables
```{r}
# check correlation for `mpg`
cor(Auto[,1:7])['mpg',]
```

Strong negative correlation with `weight`, `displacement`, `horsepower` and `cylinders`.

## 10. Boston Dataset
### (a) load dataset and overview data
```{r}
library(MASS)

dim(Boston)

summary(Boston)
```

There are 14 columns and 506 row. Columns represent the predictors and rows the observations (one per town).

### (b) Do pairplots of the data and report findings
```{r}
library(corrr)
library(magrittr)
# pairs(Boston)
Boston %>% correlate() %>% rearrange() %>% shave() %>% rplot()
```

It makes more sense to explore the data checking the correlations in this case.

The correlation matrix shows strong correlations and anti correlations in some of the predictors.

### (c) Check correlations with crime
There is a srong correlation with crime in the `rad` predictor and the `tax` predictor.

### (d) Check ranges of `crim`, `tax`, `ptratio`.
```{r}
par(mfrow=c(3,2))
hist(Boston$crim, breaks=20, main="", xlab="Crime")
plot(Boston$crim, xlab="Crime")
hist(Boston$tax, breaks=20, main="", xlab="Taxes")
plot(Boston$tax, xlab="Taxes")
hist(Boston$ptratio, breaks=20, main="", xlab="Pupil/Teacher")
plot(Boston$ptratio, xlab="Pupil/Teacher")
```

* **Crime**: There is a high range. Most have very low crime rates, but in some suburbs the crime is >= 80 with clear outliers.

* **Taxes**: There are a big chunk of suburb outliers with very high taxes (> 650).

* **Pupil/Teacher**: There is an high prevalence for a high ratio in this predictor but there aren't clear outliers.

### (e) Suburbs bound by Charles river.
There are several aproaches, but the best one in this case is to use `table`:
```{r}
table(Boston$chas)
```

Which yields 35 suburbs.

### (f) Median pupil/teacher ratio.
```{r}
median(Boston$ptratio)
```

### (g) Minimum `medv`:
```{r}
t(subset(Boston, medv == min(Boston$medv)))
summary(Boston)
```

### (h) Rooms per dwelling.
```{r}
# More than 7 rooms per dwelling
dim(subset(Boston, rm>7))[1]

# More than 8 rooms per dwelling
dim(subset(Boston, rm>8))[1]

# Compare suburbs with more than 8 rooms per dewlling on average with all.
summary(subset(Boston, rm>8))
summary(Boston)
```
